
\begin{thebibliography}{10}
\addcontentsline{toc}{part}{References}

\bibitem{BRTT93} L. Birkedal, N. Rothwell, M. Tofte and D. N. Turner. 
{\it The ML Kit, Version 1} (1993).

\bibitem{DM82} L. Damas and R. Milner.
Principal type schemes for functional programs.
{\it Proc. 9th ACM Symposium on Principles of 
Programming Languages}, pp. 207--212. ACM, New York (1982).
 
\bibitem{Kah93} S. Kahrs.
Mistakes and ambiguities in the definition of Standard ML.
Report ECS-LFCS-93-257, Univ.~of Edinburgh (1993).

\bibitem{Kah95} S. Kahrs.
Mistakes and ambiguities in the definition of Standard ML -- Addenda.
Univ.~of Edinburgh (1995).

\bibitem{bib:KST94} S. Kahrs, D. Sannella and A. Tarlecki.
{\it The Definition of Extended ML}.
Report ECS-LFCS-94-300, Univ.~of Edinburgh (1994).

\bibitem{KST93} S. Kahrs and D. Sannella and A. Tarlecki.
The semantics of Extended ML: a gentle introduction.
{\it Proc.\ Workshop on Semantics of Specification Languages},
Utrecht, 1993. Springer Workshops in Computing, 186--215 (1994).

\bibitem{MT91} R. Milner and M. Tofte.
{\it Commentary on Standard~ML}.
MIT Press (1991).

\bibitem{MTH90} R. Milner, M. Tofte and R. Harper.
{\it The Definition of Standard~ML}.
MIT Press (1990).

\bibitem{San90} D. Sannella.
Formal program development in Extended~ML for the working programmer.
{\it Proc. 3rd BCS/FACS Workshop on Refinement}, Hursley Park, 1990.
Springer Workshops in Computing, 99--130 (1991).

\bibitem{ST89} D. Sannella and A. Tarlecki.
Toward formal development of ML programs: foundations and methodology.
{\it Proc.\ Joint Conf.\ on Theory and Practice of Software Development},
Barcelona.
Springer LNCS~352, 375--389 (1989). 

\bibitem{ST91} D. Sannella and A. Tarlecki.
Extended~ML: past, present and future.
{\it Proc. 7th Workshop on Specification of Abstract Data Types}, Wusterhausen.
Springer LNCS~534, 297--322 (1991).
                                           
\end{thebibliography}