
\begin{thebibliography}{10}
\addcontentsline{toc}{section}{\refname}

%TSLS

\bibitem{Bar84} H. P. Barendregt. 
{\it The Lambda Calculus, its Syntax and Semantics.} 
2nd edition, North Holland (1984).

\bibitem{BRTT93} L. Birkedal, N. Rothwell, M. Tofte and D. N. Turner. 
{\it The ML Kit, Version 1.} (1993).

\bibitem{DM82} L. Damas and R. Milner.
Principal type schemes for functional programs.
{\it Proc.\ 9th ACM Symposium on Principles of 
Programming Languages}, pp. 207--212. ACM, New York (1982).

\bibitem{JKS96} M. Jurdzi\'nski, M. Konarski, A. Schubert.
{\it The EML Kit, Version 1.} 
Technical Report TR 96-04 (225), 
Institute of Informatics, Warsaw University (1996). 
  
\bibitem{Han93} J. Hannan.
Extended natural semantics.
{\it Journal of Functional Programming},
3(2), 123--152 (1993).
        
\bibitem{Har86} R. Harper.
{\it Introduction to Standard ML.}  
Report ECS-LFCS-86-14, Univ.~of Edinburgh (1986).

\bibitem{Kah93} S. Kahrs.
{\it Mistakes and ambiguities in the definition of Standard ML.}
Report ECS-LFCS-93-257, Univ.~of Edinburgh (1993).

\bibitem{Kah95} S. Kahrs.
{\it Mistakes and ambiguities in the definition of Standard ML -- Addenda.}
Univ.~of Edinburgh (1995).

\bibitem{Kah94} S. Kahrs.
{\it On the Static Analysis of Extended ML.}
Univ.~of Edinburgh (1994).

\bibitem{bib:KST94} S. Kahrs, D. Sannella and A. Tarlecki.
{\it The Definition of Extended ML}.
Report ECS-LFCS-94-300, Univ.~of Edinburgh (1994).

\bibitem{KST97} S. Kahrs, D. Sannella and A. Tarlecki.
{\it The Definition of Extended ML, Version 1.15 + 0.2i}.

\bibitem{KST93} S. Kahrs and D. Sannella and A. Tarlecki.
The semantics of Extended ML: a gentle introduction.
{\it Proc.\ Workshop on Semantics of Specification Languages},
Utrecht, 1993. Springer Workshops in Computing, 186--215 (1994).

\bibitem{Ler94} X. Leroy.
Manifest types, modules and separate compilation. In:
{\it Proc.\ 21st ACM Symposium on Principles of 
Programming Languages, Portland}. ACM (1994).

\bibitem{Ler92} X. Leroy.
{\it Polymorphic typing of an algorithmic language}. 
Research report 1778, INRIA (1992).

\bibitem{MT91} R. Milner and M. Tofte.
{\it Commentary on Standard~ML}.
MIT Press (1991).

\bibitem{MTH90} R. Milner, M. Tofte and R. Harper.
{\it The Definition of Standard~ML}.
MIT Press (1990).

\bibitem{R89} C. Reade.
{\it Elements of functional programming}.
Addison-Wesley (1989).

\bibitem{ST96} D. Sannella and A. Tarlecki.
Essential concepts of algebraic specification and program development. 
{\it Fundamental Aspects of Computing}, to appear.

\bibitem{ST91} D. Sannella and A. Tarlecki.
Extended~ML: past, present and future.
{\it Proc. 7th Workshop on Specification of Abstract Data Types}, Wusterhausen.
Springer LNCS~534, 297--322 (1991).

\bibitem{SW95} D. Sannella and M. Wirsing.
Specification Languages. In:
{\it Algebraic Foundations of Systems Specifications}
(E. Astesiano, H.-J. Kreowski and B. Krieg-Br\"uckner, editors).
Chapman and Hall, to appear.

\bibitem{Wir90} M. Wirsing.
Algebraic specification. In: 
{\it Handbook of Theoretical Computer Science, Vol. B}
(J. van Leeuwen, ed.), 675-788. Elsevier and MIT Press (1990).

%\bibitem{ST89} D. Sannella and A. Tarlecki.
%Toward formal development of ML programs: foundations and methodology.
%{\it Proc.\ Joint Conf.\ on Theory and Practice of Software Development},
%Barcelona. Springer LNCS~352, 375--389 (1989). 
                                 
\end{thebibliography}


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "eml-type"
%%% End: 



